The Gillespie Algorithm: Stochastic Modeling of Molecular Reaction Networks

### 2024-11-27 paper.Rmd
fixed Figure cross-references

```{r fig-dimerization-sim, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Thymine Photodimerization Simulation"}
# Reshape data for plotting
simulation_long <- melt(simulation, id.vars = "Time")
# Plot results
ggplot(simulation_long, aes(x = Time, y = value, color = variable)) +
  geom_line() +
  labs(
    title = "Dimerization Reaction Set Simulation",
    x = "Time",
    y = "Molecule Count",
    color = "Species"
  ) +
  theme_minimal()
```

@@

Both CPD and 64-PP formation follow logarithmic growth, as expected under specific UV radiation according to previously reported quantum yields (@MartinezFernandez2022; @Mouret2006). The number of reactions per unit time plateaus as the pyrimidine pairs decrease, consistent with the propensity functions.
Due to the stochastic nature of molecular reactions, this algorithm reflects fluctuations of photodimerization as it models the probabilistic process of each reaction.

@@

The simulation results are presented in two figures, each offering unique insights into the behavior of the decay process. Figure \ref(fig-decay-linear) displays the decay on a linear scale, illustrating the overall exponential decline of the radioactive substance over time. Figure \ref(fig-decay-log) presents the same data on a logarithmic scale, highlighting the exponential nature of the process by linearizing the decay trajectory.

@@

#### Analysis of Figure \ref(fig-decay-linear)
Figure \ref{fig:fig-decay-linear} illustrates the exponential decay of \( R \) over time. The steep initial decline in the curve corresponds to the high decay rate at the beginning of the simulation when \( R \) is at its maximum. As \( R \) decreases, the curve flattens, reflecting the slower decay rate. This behavior exemplifies the exponential decay dynamics, where the number of remaining molecules reduces rapidly at first and progressively lessens as the substance depletes.

@@

```{r fig-decay-linear, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Radioactive Decay Simulation in Linear Scale"}

@@

\newpage
#### Analysis of Figure \ref(fig-decay-log)
 Figure \ref(fig-decay-log) depicts the same decay trajectory on a logarithmic scale, offering a complementary perspective. On this scale, the exponential decay law is linearized, and the trajectory appears as a straight line. The slope of the line corresponds to the negative decay constant, \(-c\), providing a quantitative validation of the theoretical model.

@@

```{r fig-decay-log, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Radioactive Decay Simulation in Logarithmic Scale"}
# Logarithmic-scale plot
ggplot(simulation, aes(x = Time, y = R)) +
  geom_line(color = "red") +
  scale_y_log10() +
  labs(
    x = "Time",
    y = "Remaining Radioactive Substance (log(R))"
  ) +
  theme_minimal()

>>>>

```{r fig-dimerization-sim, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Thymine Photodimerization Reaction Set Simulation \\label{fig:dimerization-sim}"}
# Reshape data for plotting
simulation_long <- melt(simulation, id.vars = "Time")
# Plot results
ggplot(simulation_long, aes(x = Time, y = value, color = variable)) +
  geom_line() +
  labs(
    x = "Time",
    y = "Molecule Count",
    color = "Species"
  ) +
  theme_minimal()
```

@@

Both CPD and 64-PP formation follow logarithmic growth, as expected under specific UV radiation according to previously reported quantum yields (@MartinezFernandez2022; @Mouret2006). The number of reactions per unit time plateaus as the pyrimidine pairs decrease, consistent with the propensity functions.
Due to the stochastic nature of molecular reactions, this algorithm reflects fluctuations of photodimerization as it models the probabilistic process of each reaction (Figure \ref{fig:dimerization-sim}).

@@

The simulation results are presented in two figures, each offering unique insights into the behavior of the decay process. Figure \ref{fig:decay-linear} displays the decay on a linear scale, illustrating the overall exponential decline of the radioactive substance over time. Figure \ref{fig:decay-log} presents the same data on a logarithmic scale, highlighting the exponential nature of the process by linearizing the decay trajectory.

@@

#### Analysis of Figure \ref{fig:decay-linear}
Figure \ref{fig:decay-linear} illustrates the exponential decay of \( R \) over time. The steep initial decline in the curve corresponds to the high decay rate at the beginning of the simulation when \( R \) is at its maximum. As \( R \) decreases, the curve flattens, reflecting the slower decay rate. This behavior exemplifies the exponential decay dynamics, where the number of remaining molecules reduces rapidly at first and progressively lessens as the substance depletes.

@@

```{r fig-decay-linear, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Radioactive Decay Simulation in Linear Scale \\label{fig:decay-linear}"}

@@

\newpage
#### Analysis of Figure \ref{fig:decay-log}
 Figure \ref{fig:decay-log} depicts the same decay trajectory on a logarithmic scale, offering a complementary perspective. On this scale, the exponential decay law is linearized, and the trajectory appears as a straight line. The slope of the line corresponds to the negative decay constant, \(-c\), providing a quantitative validation of the theoretical model.

@@

```{r fig-decay-log, echo=FALSE, warning=FALSE, message=FALSE, fig.cap="Radioactive Decay Simulation in Logarithmic Scale \\label{fig:decay-log}"}
# Logarithmic-scale plot
ggplot(simulation, aes(x = Time, y = R)) +
  geom_line(color = "red") +
  scale_y_log10() +
  labs(
    x = "Time",
    y = "Remaining Radioactive Substance (log(R))"
  ) +
  theme_minimal()
